---
title: "Global Salaries"
author: "Dharshan Lakshmikanthan"
execute: 
  echo: false
---

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(brms)
library(ggplot2)
library(gtsummary)
library(tidybayes)
library(primer.data)

# Load and clean data
salary_data <- read_csv('Salary_Data_Based_country_and_race.csv')

# Clean column names by replacing spaces with underscores
salary_data <- salary_data %>%
  rename_with(~str_replace_all(., " ", "_")) %>%
  mutate(
    education = as.factor(Education_Level),
    experience = as.numeric(Years_of_Experience)
  ) %>%
  drop_na()

# Print column names to verify
print(colnames(salary_data))
```


```{r}
#| cache: true
#| warning: false
# Fit the model
fit_salary <- brm(formula = Salary ~ experience + education,
    data = salary_data,
    family = gaussian(),
    refresh = 0,
    silent = 2,
    seed = 9)
```

```{r}
#| label: regression_table
#| cache: false
tbl_regression(fit_salary)

"This table summarizes the model fit for the global salary data."
```

```{r}
#| cache: true
# Prepare new data for predictions
ndata <- expand_grid(
  experience = c(5, 10),
  education = levels(salary_data$education)[1:2]  # Assuming you want the first two levels
)

# Generate posterior predictions
posterior_samples <- posterior_predict(fit_salary, newdata = ndata)

# Convert predictions to a data frame
predictions_df <- as_tibble(posterior_samples) %>%
  mutate(draw = row_number()) %>%
  pivot_longer(-draw, names_to = "scenario", values_to = "prediction") %>%
  mutate(
    experience = rep(ndata$experience, each = nrow(posterior_samples)),
    education = rep(ndata$education, each = nrow(posterior_samples))
  )

# Calculate mean predictions and effects
mean_predictions <- predictions_df %>%
  group_by(experience, education) %>%
  summarise(mean_prediction = mean(prediction), .groups = 'drop') %>%
  pivot_wider(names_from = education, values_from = mean_prediction) %>%
  mutate(
    education_effect = .[[3]] - .[[2]],
    experience_effect = (.[[2]][experience == 10] - .[[2]][experience == 5])
  ) %>%
  pivot_longer(cols = c(education_effect, experience_effect),
               names_to = "effect_type",
               values_to = "effect")
```

```{r}
#| cache: false
# Plot predictions
ggplot(mean_predictions, aes(x = effect, fill = effect_type)) +
  geom_histogram(bins = 100, aes(y = after_stat(count / sum(count)))) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_classic() +
  labs(title = "Posterior Distribution of Effects on Salary",
       x = "Difference in Salary",
       y = "Probability",
       fill = "Effect Type")
```


We used global salary data to measure the causal effects of education level and years of experience on salaries across various countries and industries in 2024. We modeled Salary as a linear function of Years of Experience and Education Level. The measures of uncertainty around these causal effects are reflected in the distributions of the estimated effects, which we visualize using histograms of the effect draws to assess their variability.

